Empirical normative theories

### 2020-06-24 simple.tex
Further clarification of notations and improvements

Given $i \in I, \phi \in \Phi, \hist \in \allhist$, say that $\hist$ leads strongly to $\phi$ iff $\hist \ileadstoi \phi \land \hist \nileadstoi ¬\phi$.
Define $\hist \ileadstosti \phi$ iff $[\hist \text{ leads strongly to } \phi] \land [\phi = \propsure ⇒ \hist \text{ leads strongly to } \propposs]$.
Given $\hist \in \allhist$, let $\hist_\mathit{end}$ denote the set containing the last two elements of the sequence $\hist$ if the sequence has at least two elements; the singleton set containing the last element of the sequence if it has exactly one element; and the empty set if $\hist = \emptyset$.
Given $i \in I, \phi \in \Phi, \ar \in \allargs$, define $\ar \ileadstosti \phi$ iff $\forall \hist \in \allhist \suchthat \ar \in \hist_\mathit{end}: \hist \ileadstosti \phi$.

@@

\section{Additive single-answer empirical theories}
\subsection{Definition}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
${\gbeatsinv}(\allargs)$ finite,
and $+$ a commutative associative internal binary operation on $\allargs$. Given a finite $\args \subseteq \allargs$, the notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument. 
\commentOC{Check about finiteness.}
\commentOC{${\gbeatsinv}$ need not be finite, only those even ones: the transitive closure of ${\gleadstoinv}(\Phi)$ through ${\gbeatsinv}^2$, that is, $({\gbeatsinv})^2[{\gleadstoinv}(\Phi)] = \cup_{k \in 2\N}({\gbeatsinv}^{k}({\gleadstoinv}(\Phi)))$.}
\commentOC{I should only define $\ar \in \allargs$ plus $\ard$.}

This defines a function: $\forall \ar_1 \gbeats \ar_0: f(\ar_1, \ar_0) = \ar_0 + {\gbeatsinv}(\ar_1) = \set{\ar_0 + \ard_2 \suchthat \ard_2 \in {\gbeatsinv}(\ar_1)}$, and $f(\ar_1, \ar_0) = \emptyset$ otherwise. Thus, $f = \set{((\ar_1, \ar_0), \ar_2) \suchthat \exists \ard_2 \in \allargs \suchthat \ard_2 \gbeats \ar_1 \gbeats \ar_0 \land \ar_2 = \ar_0 + \ard_2}$.
Also, from the previous definition, $g(\ar_0)(\ar_0 + \argsd_2) = dom(f^{-1}(\ar_0 + \argsd_2) \cap (\allargs × \set{\ar_0})) = {\gbeats}(\argsd_2) \cap {\gbeatsinv}(\ar_0)$ because $(\ar_1, \ar_0) \in f^{-1}(\ar_0 + \argsd_2) ⇔ \exists \ar_2 \in \ar_0 + \argsd_2 \suchthat \ar_2 \in f(\ar_1, \ar_0) ⇔ \ar_1 \gbeats \ar_0 \land \exists \ard \in \argsd_2 \suchthat \ard \gbeats \ar_1$.
\commentOC{Could be false. Define }

\subsection{Falsifiability}
\begin{proposition}
	An empirical theory has only finite sets $\gargs[i, \phi]$, and an upper bound on their size exists that is independent of ${\ileadsto}$.
\end{proposition}
\begin{proof}
	Given $i, \phi$. Observe that $\forall \ard: d^{i, \phi}(\ard) \subseteq {\gbeatsmt}(\ard)$. Thus, $d^{i, \phi}$ is finite. Also, the transitive closure through $\sigma^i$ changes a finite number of time, otherwise it is possible to build an infinite chain in ${\gbeatsmt}$.
\end{proof}

Given $i \in I, \phi \in \Phi, \ar_1, \ar_0 \in \allargs$, define $\ar_1 \ibeatse[!\phi] \ar_0$ iff $\exists \hist \in \allhist \suchthat \hist \text{ ends with } \set{\ar_1, \ar_0} \land \hist \nileadstosts \phi$.

\begin{proposition}
	$\forall i \in I, \phi \in \Phi, \ar_0 \in \allargs$:
\begin{equation}
	\label{eq:st-in-terms-of-e}
	\ar_0 \nileadstosts \phi ⇔ [\exists \ar_1 \in \allargs \suchthat \ar_1 \ibeatse[!\phi] \ar_0].
\end{equation}
\end{proposition}
\begin{proof}
	By definition of ${\ileadstosts}$, $\ar_0 \ileadstosts \phi ⇔ \forall \hist \in \allhist \suchthat [\exists \ar_1 \in \allargs \suchthat \hist \text{ ends with } \set{\ar_1, \ar_0}]: \hist \ileadstosts \phi.$ Equivalently, $\ar_0 \nileadstosts \phi ⇔ \exists \hist \in \allhist \suchthat [\exists \ar_1 \suchthat \hist \text{ ends with } \set{\ar_1, \ar_0}] \land \hist \nileadstosts \phi$. This is also equivalent to $\ar_0 \nileadstosts \phi ⇔ \exists \hist \in \allhist, \ar_1 \in \allargs \suchthat \hist \text{ ends with } \set{\ar_1, \ar_0} \land \hist \nileadstosts \phi$. By definition of ${\ibeatse}$, this is the proposition to be proven.
\end{proof}

@@

Given $i \in I, \phi \in {\gleadsto}(\allargs)$, define the $\phi$-protectors of $\ard$ as $d^{i, \phi}(\ard) = [{\gbeatsinv} \circ ({\gbeatsinv} \cap {\ibeatseinv[!\phi]})](\ard)$.
Define the successors of $(\phi, \ar, \ard)$ as $\sigma^i(\phi, \ar, \ard) = \set{\phi} × \set{\ar + \ard} × d^{i, \phi}(\ard)$. Given $i \in I, \phi \in {\gleadsto}(\allargs)$, define $\gargs[i, \phi]$ as the transitive closure of $(\phi, \emptyset, \ar)$ under $\sigma^i$, with $\ar \gleadsto \phi$.
\commentOC{Assuming only one $\ar$ leads to that $\phi$.}
Define $\gargs[i, \phi] = \bigcup_{k \in 2\N} \gargs[i, \phi, k]$, and $\gargs = \cup_{i \in I, \phi \in {\gleadsto}(\allargs)} \gargs[i, \phi]$.

@@

\section{Empirical theories}
\subsection{Falsifiability}
\commentOC{I want: An empirical theory for a normative theory is something like a normative and falsifiable claim. It is also configurable: it must be possible to guarantee that the falsification be easy (if the claim is false) depending on the kind of claim.}

>>>>

Given $i \in I, \phi \in \Phi, \hist \in \allhist$, say that $\hist$ leads strongly to $\phi$ iff $\hist \ileadstoi \phi \land \hist \nileadstoi ¬\phi$.
Define $\hist \ileadstosti \phi$ iff $[\hist \text{ leads strongly to } \phi] \land [\phi = \propsure ⇒ \hist \text{ leads strongly to } \propposs]$.
Given $\hist \in \allhist$, let $\histend$ denote the set containing the last two elements of the sequence $\hist$ if the sequence has at least two elements; the singleton set containing the last element of the sequence if it has exactly one element; and the empty set if $\hist = \emptyset$.
Given $i \in I, \phi \in \Phi, \ar \in \allargs$, define $\ar \ileadstosti \phi$ iff $\forall \hist \in \allhist \suchthat \ar \in \histend: \hist \ileadstosti \phi$.

@@

\section{Additive single-answer empirical theories}
\subsection{Definition}
\commentOC{I leave the definition as is, but for the moment I only consider ${\gleadsto}$, a relation that does not depend on $i$; and I do not use $q(i)$.}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, \alldargs, \dbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq (\allargs × \Phi) \cap (\alldargs × \Phi)$, 
${\gbeats} \subseteq \allargs × \allargs$,
${\dbeats} \subseteq \alldargs × \allargs$ and is injective (thus $\forall \ar \in \allargs: {\dbeatsinv}(\ar)$ is a singleton or empty). 
Define $\df \subseteq \allargs × \alldargs$ as $\df = {\dbeatsinv} \circ {\gbeatsinv}$. 
The symbol $+$ in an empirical theory stands for a function $+: \bigcup_{\ar \in \allargs}(\set{s} × \df(\ar)) → \allargs$.
As usual, $\ar + \ard$ means $+(\ar, \ard)$. 
Define $\dfp \subseteq \allargs × \allargs$ as, $\forall \ar \in \allargs$: $\dfp(\ar) = \ar + \df(\ar) = \set{\ar + \ard \suchthat \ard \in \df(\ar)}$.
An empirical theory must satisfy
$\forall \ar \in {\gleadsto}(\allargs): \exists k \in \N^* \suchthat \dfp^k(\ar) = \emptyset$.

\subsection{Falsifiability}
Given $i \in I, \phi \in \Phi, \ar_1, \ar_0 \in \allargs$, define $\ar_1 \ibeatsi \ar_0$ iff $\exists \hist \in \allhist \suchthat \set{\ar_1, \ar_0} \subseteq \histend \land \hist \nileadstosti \phi$.

\begin{proposition}
	$\forall i \in I, \phi \in \Phi, \ar_0 \in \allargs$:
\begin{equation}
	\label{eq:st-in-terms-of-e}
	\ar_0 \nileadstosti \phi ⇔ [\exists \ar_1 \in \allargs \suchthat \ar_1 \ibeatsi \ar_0].
\end{equation}
\end{proposition}
\begin{proof}
	By definition of ${\ileadstosti}$, $\ar_0 \ileadstosti \phi ⇔ \forall \hist \in \allhist \suchthat \ar_0 \in \histend: \hist \ileadstosti \phi$. 
Equivalently, $\ar_0 \ileadstosti \phi ⇔ \forall \hist \in \allhist \suchthat [\exists \ar_1 \in \allargs \suchthat \set{\ar_1, \ar_0} \subseteq \histend]: \hist \ileadstosti \phi$. 
Still equivalently, $\ar_0 \nileadstosti \phi ⇔ \exists \hist \in \allhist \suchthat [\exists \ar_1 \in \allargs \suchthat \set{\ar_1, \ar_0} \subseteq \histend] \land \hist \nileadstosti \phi$. 
This can also be written: $\ar_0 \nileadstosti \phi ⇔ \exists \hist \in \allhist, \ar_1 \in \allargs \suchthat \set{\ar_1, \ar_0} \subseteq \histend \land \hist \nileadstosti \phi$. By definition of ${\ibeatsi}$, this is the proposition to be proven.
\end{proof}

Given $i \in I, \phi \in {\gleadsto}(\allargs), \ar \in \allargs$, define the \emph{$\phi$-defensors of $\ar$ required for $i$}, $d^{i, \phi}(\ar) \subseteq \alldargs$, as $d^{i, \phi}(\ar) = [{\dbeatsinv} \circ ({\gbeatsinv} \cap {\ibeatsiinv})](\ar)$.
%Given $i \in I, \phi \in {\gleadsto}(\allargs), \ar \in \allargs, \ard \in \df(\ar)$, define the \emph{successors} of $(\ar, \ard)$ as $\sigma^{i, \phi}(\ar, \ard) = \set{\ar + \ard} × d^{i, \phi}(\ar + \ard)$. 
Given $i \in I, \phi \in {\gleadsto}(\allargs)$, define $\gargs[i, \phi] \subseteq \allargs$ as the transitive closure of ${\gleadstoinv}(\phi)$ under $\dfp$: $\gargs[i, \phi] = \cup_{k \in \N} \dfp^k({\gleadstoinv}(\phi))$, where $\dfp^0$ is the identity function.

%\begin{proposition}
%	An empirical theory has only finite sets $\gargs[i, \phi]$, and an upper bound on their size exists that is independent of ${\ileadsto}$.
%\end{proposition}
%\begin{proof}
%	Given $i, \phi$. Observe that $\forall \ard: d^{i, \phi}(\ard) \subseteq {\gbeatsmt}(\ard)$. Thus, $d^{i, \phi}$ is finite. Also, the transitive closure through $\sigma^i$ changes a finite number of time, otherwise it is possible to build an infinite chain in ${\gbeatsmt}$.
%\end{proof}

@@

@@

An additive theory need not have freedom for defining planned attacks on additive terms: this should be a consequence on the planned attacks on defenses and on initial arguments only. This makes the theory easier to define, as only attacks on partial arguments are to be defined.

\section{Empirical theories}
\subsection{Simple to complex}
From an additive single-answer empirical theory, we can define a function: $\forall \ar_1 \gbeats \ar_0: f(\ar_1, \ar_0) = \ar_0 + {\gbeatsinv}(\ar_1) = \set{\ar_0 + \ard_2 \suchthat \ard_2 \in {\gbeatsinv}(\ar_1)}$, and $f(\ar_1, \ar_0) = \emptyset$ otherwise. Thus, $f = \set{((\ar_1, \ar_0), \ar_2) \suchthat \exists \ard_2 \in \allargs \suchthat \ard_2 \gbeats \ar_1 \gbeats \ar_0 \land \ar_2 = \ar_0 + \ard_2}$.
Also, from the previous definition, $g(\ar_0)(\ar_0 + \argsd_2) = dom(f^{-1}(\ar_0 + \argsd_2) \cap (\allargs × \set{\ar_0})) = {\gbeats}(\argsd_2) \cap {\gbeatsinv}(\ar_0)$ because $(\ar_1, \ar_0) \in f^{-1}(\ar_0 + \argsd_2) ⇔ \exists \ar_2 \in \ar_0 + \argsd_2 \suchthat \ar_2 \in f(\ar_1, \ar_0) ⇔ \ar_1 \gbeats \ar_0 \land \exists \ard \in \argsd_2 \suchthat \ard \gbeats \ar_1$.
\commentOC{Could be false. Define }

\subsection{Falsifiability}
\commentOC{I want: An empirical theory for a normative theory is something like a normative and falsifiable claim. It is also configurable: it must be possible to guarantee that the falsification be easy (if the claim is false) depending on the kind of claim.}
